---
title: CV
name: Grace
surname: Kumaishi
position: "Master's student"
address: "Bren School of Environmental Science and Management, UC Santa Barbara"
phone: +1 831-212-2293
email: "gkumaishi@ucsb.edu"
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Marie is a Polish and naturalized-French physicist and chemist who conducts pioneering research on radioactivity."
output: 
  vitae::markdowncv:
    theme: davewhipp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

```{r echo = FALSE, message=FALSE, fig.align='center', fig.pos='H', out.width = "800px", out.height = "1100px"}
knitr::include_graphics("./CV/cv.pdf")
```

## Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bren School of Environmental Science and Management, UCSB", "Expected 2022", "Santa Barbara, CA", "Master of Environmental Science and Management,",
  "University of California, Los Angeles (UCLA)", "2013-2017", "Los Angeles, CA", "Bachelor of Science in Ecology, Behavior, and Evolution,"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Research Experience

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Common Murre Restoration Project, Humboldt State University", "4/2020 - 8/2020", "Pacifica, CA", "Wildlife Technician,",
  "Point Blue Conservation Science", "4/2019 - 3/2020", "Petaluma, CA", "Farallon Program Research Intern,",
  "Point Blue Conservation Science", "9/2017 - 1/2019", "Petaluma, CA", "Marine Laboratory Research Assistant,",
    "UCLA Department of Ecology, Behavior, and Evolution", "7/2017", "Los Angeles, CA", "Laboratory Research Assistant,",
    "UCLA Department of Ecology, Behavior, and Evolution", "4/2016 - 3/2017", "Los Angeles, CA", "Undergraduate Research Assistant,"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Volunteer Experience

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Oceans 360, Wildlife Conservation Network Expo", "10/2018 - 10/2019", "San Francisco, CA", "Interpreter,",
  "Surin Project", "8/2014", "Surin, Thailand", "Sanctuary Volunteer,",
  "Seymour Center at Long Marine Lab, UCSC", "3/2011 - 4/2013", "Santa Cruz, CA", "Exhibit Guide,"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Presentations

**Grace Kumaishi**, Pete Warzybok, Michael E. Johns. (2020) Pacific Seabird Group 47th Annual Meeting; Portland, Oregon. **Oral**. Associations between annual changes in chick diet and breeding success of pigeon guillemots on Southeast Farallon Island. 

## Skills and Certifications

**Software:** R, R Studio, QGIS, ImageJ, Microsoft Office Suite

**Certifications:** Open Water Diver (PADI), First Aid & CPR (American Heart Association)

**Languages:** English, Conversational Spanish
