---
title: "Exploratory report"
description: |
  using juvenile snowshoe hare statistics 
author:
  - name: Grace Kumaishi
date: 11-15-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Install packages:
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
library(ggbeeswarm)
library(ggplot2)
library(effsize)
library(broom)
```

```{r, include=FALSE}
# Read in data:
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv")) %>%
  clean_names()
```

## Introduction
This report explores physical data collected from trapped snowshoe hares (*Lepus americanus*) in the Bonanza Creek Experimental Forest between 1999 and 2012. The Bonanza Creek Experimental Forest is located in Tanana Valley, Alaska, and consists of numerous habitats that vary in density of cover (Feierabend and Kielland 2014). This report focuses on three sites of interest: Bonanza Riparian (a flood zone consisting of deciduous trees), Bonanza Mature (a mature site consisting of both deciduous trees and black spruce, *Picea mariana*), and Bonanza Black Spruce (a site used for live-trapping located in a mature black spruce stand) (Kielland et al. 2017). The analyses included here focus specifically on the recorded weight of juvenile hares trapped over the course of the study. In addition to exploratory visualizations of weight comparisons between sites and sex, we explore the relationship between juvenile weight and hind foot length. 

## Data and analyses

Snowshoe hare measurements were collected and provided by Dr. Knut Kielland and collegues at the Bonanza Creek Long-Term Ecological Research (LTER) site located in Tanana Valley, Alaska. The data contain observations of adult and juvenile snowshoe hares trapped between 1999 and 2012. In addition to exploratory data visualization, weight measurements of juvenile male and female snowshoe hares are compared using two-sample t-tests (significance level ($\alpha$) of 0.05), while differences between sexes are described with Cohen’s *d* effect size. The relationship between juvenile hare weight and hind foot length is explored using a simple linear regression. Analyses were performed in R version 4.0.2 using RStudio version 1.3.1093.

## Exploratory findings

### 1) Annual juvenile hare trap counts

```{r}
# Create data set including only juvenile hares captured, create column for year:
juv_hares <- bonanza_hares %>% 
  filter(age == "j") %>% 
  mutate(date_new = mdy(date)) %>%
  mutate(year_new = year(date_new)) %>% 
  count(year_new)

# Find the mean and median number of juvenile hares trapped per year:
juv_annual_stats <- juv_hares %>% 
  summarize(mean_count = mean(n),
            median_count = median(n))
```

The maximum number of juvenile hares trapped during this study occurred in the first two years (1999 and 2000) with `r juv_hares[1,2]` and `r juv_hares[2,2]` individuals trapped, respectively. The following years fluctuate between zero (this occurred during years in which the age of hares was not specified in the data) and just under 50, which corroborates the authors' statement that hare populations fluctuate over 8 to 11 years (Kielland et al. 2017). On average, `r juv_annual_stats[1,1]` juvenile hares were captured per year, with a median value of `r juv_annual_stats[1,2]`. In order to standardize the juvenile hare population in future analyses, trap counts could be divided by time or trapping effort.

```{r, fig.align = "center"}
# Data visualization of number of juv hares trapped per year:
ggplot(data = juv_hares, aes(x = year_new, y = n)) +
  geom_col(fill = "royalblue") +
  labs(x = "Year", y = "Count") +
  theme_minimal() +
  scale_x_continuous(breaks = c(1999, 2001, 2003, 2005, 2007, 2009, 2011))

# Zeros for 2002 and 2009--data was still collected in these years but age was not specified
```
**Figure 1:** Number of juvenile hares trapped between 1999 and 2012 in the Bonanza Creek Experimental Forest, Tanana Valley, Alaska. Data: Kielland et al. (2017). 


### 2) Visualize juvenile hare weights

```{r, fig.align = "center"}

# Create a data set of juvenile hares including sex, site, and weight:
juv_hares_sex_site <- bonanza_hares %>% 
  filter(age == "j") %>% 
  select("sex", "grid", "weight") %>% 
  mutate(grid = case_when( # Rename sites in dataset so that the full name shows up in visualizations
    grid == "bonbs" ~ "Bonanza Black Spruce", 
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian"))

# Create Vectors for male and female hares:
juv_hares_f <- juv_hares_sex_site %>% 
  filter(sex == "f") %>% 
  pull(weight)

juv_hares_m <- juv_hares_sex_site %>% 
  filter(sex == "m") %>% 
  pull(weight)

# Create Vectors for each site:
juv_hares_bonrip <- juv_hares_sex_site %>% 
  filter(grid == "Bonanza Riparian") %>% 
  pull(weight)

juv_hares_bonmat <- juv_hares_sex_site %>% 
  filter(grid == "Bonanza Mature") %>% 
  pull(weight)

juv_hares_bonbs <- juv_hares_sex_site %>% 
  filter(grid == "Bonanza Black Spruce") %>% 
  pull(weight)

# Get the means and SD's for each (make sure to remove NA's from data!!!):
male_mean <- mean(juv_hares_m)
female_mean <- mean(juv_hares_f, na.rm = TRUE)
male_sd <- sd(juv_hares_m)
female_sd <- sd(juv_hares_f, na.rm = TRUE)
bonrip_mean <- mean(juv_hares_bonrip, na.rm = TRUE)
bonrip_sd <- sd(juv_hares_bonrip, na.rm = TRUE)
bonmat_mean <- mean(juv_hares_bonmat, na.rm = TRUE)
bonmat_sd <- sd(juv_hares_bonmat, na.rm = TRUE)
bonbs_mean <- mean(juv_hares_bonbs, na.rm = TRUE)
bonbs_sd <- sd(juv_hares_bonbs, na.rm = TRUE)
```

To begin visualizing juvenile hare weights, a density plot including both sex and site was created (Figure 2). This plot allows us to visualize the spread of weights for male, female and unknown juvenile hares between each of the three sites. The density plot shows female weight peaking earlier than male weight in the Bonanza Mature and Bonanza Riparian sites, indicating that females may be smaller on average here. These trends are explored further in Figures 3 and 4, which allow us to examine the relationship between weight, sex, and site individually. When comparing weight and sex alone, male hares are heavier on average than female hares (with mean weights of `r round(male_mean,2)` g and `r round(female_mean,2)` g respectively). Over the three different sites, juvenile hares trapped at Bonanza Black Spruce weighed the most on average (`r round(bonbs_mean,2)` g), while the weights of juvenile hares captured at Bonanza Mature and Bonanza Riparian were relatively similar, with averages of `r round(bonmat_mean,2)` g and `r round(bonrip_mean,2)` g, respectively. Distributions of weight for both sex and site are relatively symmetric and lack notable outliers. 

```{r, fig.align = "center"}
# Visualization looking at both sex, site and weight together:
ggplot(data = juv_hares_sex_site, aes(x = weight)) +
  geom_density(aes(fill = sex), alpha = 0.5) + # Keep alpha outside aes() so that it doesn't show up in legend!!
  facet_wrap(~grid, nrow = 3) +
  labs(x = "Weight (g)", y = "Density") +
  theme_minimal() 

# This graph is okay, but a bit difficult to comprehend what is happening between the sexes. We can see that peaks of males tend to be further to the right than the peaks of females, indicating that males are heavier. I decided to break up the analysis into two separate graphs--one for "sex" and one for "site" so that I could look at these relationships on their own. I could have used a beeswarm plot for all three visualizations, but thought it would be more interesting to try something different. 
```

**Figure 2:** Density plots comparing weight in grams between sexes of juvenile snowshoe hares trapped at three different sites: Bonanza Black Spruce, Bonanza Mature, and Bonanza Riparian in the Bonanza Creek Experimental Forest, Tanana Valley, Alaska. Data: Kielland et al. (2017).

```{r, fig.align = "center"}
# Create data set looking at just sex and weight to add mean to each plot:
juv_sex_stats <- juv_hares_sex_site %>% 
  group_by(sex) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE))

# Visualization (sex and weight) using both beeswarm and boxplot:
ggplot() +
  geom_beeswarm(data = juv_hares_sex_site, 
                aes(x = sex, y = weight, color = sex),
                size = 2,
                alpha = 0.5) +
  geom_boxplot(data = juv_hares_sex_site, 
             aes(x = sex, y = weight),
             width = 0.3,
             size = 0.5,
             alpha = 0) +
  geom_point(data = juv_sex_stats, # Add mean
             aes(x = sex, y = mean_weight)) +
  theme_minimal() +
  labs(x = "Sex", y = "Weight (g)") +
   theme(legend.position = "none")
```

**Figure 3:** Weight observations (g) for juvenile snowshoe hares separated by sex (female, male, and NA). Pink (female), blue (male), and gray (sex not recorded) points indicate individual observations for weight (grams) of juvenile hares that were trapped between 1999 and 2012. Box endpoints indicate the 25^th^ and 75^th^ percentile values. The median value is indicated by a black line and the mean value is indicated by a black dot within the boxplot. Data: Kielland et al. (2017). 

```{r, fig.align = "center"}
# Create data set looking at just site and weight to add mean to each plot:
juv_site_stats <- juv_hares_sex_site %>% 
  group_by(grid) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE))

# Visualization (site and weight) using both beeswarm and boxplot:
ggplot() +
  geom_beeswarm(data = juv_hares_sex_site, 
                aes(x = grid, y = weight, color = grid),
                size = 2,
                alpha = 0.5) +
  geom_boxplot(data = juv_hares_sex_site, 
             aes(x = grid, y = weight),
             width = 0.3,
             size = 0.5,
             alpha = 0) +
    geom_point(data = juv_site_stats, # Add mean
             aes(x = grid, y = mean_weight)) +
  theme_minimal() +
  labs(x = "Site", y = "Weight (g)") +
  theme(legend.position = "none")
```

**Figure 4:** Weight observations (g) for juvenile snowshoe hares separated at three different sites in Tanana Valley, Alaska. Pink (Bonanza Black Spruce), green (Bonanza Mature), and blue (Bonanza Riparian) points indicate individual observations for weight (grams) of juvenile hares that were trapped between 1999 and 2012. Box endpoints indicate the 25^th^ and 75^th^ percentile values. The median value is indicated by a black line and the mean value is indicated by a black dot within the boxplot. Data: Kielland et al. (2017). 

### 3) Juvenile weight comparison (male & female snowshoe hares)

```{r, include = FALSE}
# Plot histograms and QQ-plots to check assumptions:
# Histogram
ggplot(data = juv_hares_sex_site, aes(x = weight)) +
  geom_histogram(bins = 15) +
  facet_wrap(~sex) +
  theme_minimal() # appears relatively normal, males look less normal than females, perhaps negatively skewed

# QQ-plot
ggplot(data = juv_hares_sex_site, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex) +
  theme_minimal() # appears relatively linear, males are less linear than females, perhaps negatively skewed

# The histogram and QQ-plots suggest that assumptions of normality hold, & sample size additionally > Central Limit Theorem (means will be normally distributed). Will proceed with two-sample t-test. 

# Run the t-test:
juv_hares_ttest <- t.test(juv_hares_f, juv_hares_m)

# Get the tidy model results to call outputs in-line:
juv_hares_ttest_tidy <- tidy(juv_hares_ttest)

# Get the effect size:
juv_hares_effsize <- cohen.d(juv_hares_f, juv_hares_m, na.rm = TRUE)
```

On average, juvenile male snowshoe hares weigh more than juvenile female snowshoe hares (`r round(male_mean,2)` $\pm$ `r round(male_sd, 2)` and `r round(female_mean,2)` $\pm$ `r round(female_sd, 2)` mm, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r round(male_mean, 2) - round(female_mean, 2)` mm (a `r round((male_mean - female_mean)/((male_mean + female_mean)/2)*100, 2)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(juv_hares_ttest_tidy$parameter,2)`) = `r round(juv_hares_ttest_tidy$statistic,2)`, p < 0.01), and the effect size is small (Cohen's *d* = `r round(juv_hares_effsize$estimate,2)`).

**Table 1:** Descriptive statistics (mean, standard deviation, and sample size) of juvenile hares captured between 1999 and 2012. Data: Kielland et al. (2017).

```{r}
# Create data table
juv_hare_stats <- juv_hares_sex_site %>% 
  group_by(sex) %>% 
  summarize(mean_weight = round(mean(weight, na.rm = TRUE),2),
            sd_weight = round(sd(weight, na.rm = TRUE),2),
            n = n()) %>% 
  mutate(sex = case_when( 
    sex == "f" ~ "Female", 
    sex == "m" ~ "Male"))

juv_hare_stats %>% 
  kable(col.names = c("Sex", "Mean weight (g)", "Standard deviation (g)", "Sample size")) %>% 
  kable_minimal()
```

### 4) Relationship between juvenile weight & hind foot length

```{r}
# Create data set looking at juvenile hare weight and hind foot length
juv_hares_weight_foot <- bonanza_hares %>% 
  filter(age == "j") %>% 
  select("weight", "hindft")
```

```{r, include = FALSE}
# Linear model, stored as hares_lm:
hares_lm <- lm(weight ~ hindft, data = juv_hares_weight_foot)

# Return the complete overview:
summary(hares_lm)

# Get tidy versions of the model output to call later on in text
hares_lm_tidy <- tidy(hares_lm)
hares_lm_glance <- glance(hares_lm)

# See model diagnostics:
plot(hares_lm) # Assumptions for a linear model are moderately met. However, presence of outliers in Residuals vs. Leverage plot may pose some concern.

# Get Pearson's r correlation
hares_cor <- cor.test(juv_hares_weight_foot$weight, juv_hares_weight_foot$hindft)

hares_cor

#Tidy version of correlation output:
hares_cor_tidy <- tidy(hares_cor)
```

The linear relationship between juvenile hare weight and hind foot length appears somewhat linear and positively correlated (Figure 5). Simple linear regression revealed that weight moderately predicts juvenile hare hind foot length, (p < 0.001, R^2^ = `r round(hares_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(hares_lm_tidy$estimate[2], 2)` cm (i.e., for each one centimeter increase in weight we expect an average increase in hind foot length of `r round(hares_lm_tidy$estimate[2], 2)` cm). Hind foot length and weight are moderately, and significantly, positively correlated (Pearson's *r* = `r round(hares_cor_tidy$estimate,2)`, p < 0.001). Diagnostic plots (hidden from output) reveal relatively normally distributed and homoscedastic residuals. 

Although we proceeded with a linear model, our assumptions were moderately met and suggest limited reason for concern. The R^2^ value of `r round(hares_lm_glance$r.squared,2)` indicates that about 30% of variance in hind foot length can be predicted by weight. The output graph showing Residuals vs Leverage contains a couple of labeled points, indicating that there may be outliers in the data that are influencing this model. 

```{r, fig.align = "center"}
# Create scatterplot showing data for juvenile hares weight versus hind foot length. Add a trendline to data. 
ggplot(data = juv_hares_weight_foot, aes(x = weight, y = hindft)) + 
  geom_point(color = "royalblue") +
  geom_smooth(method = lm, # This is how you add trendline
              se = FALSE, # Remove confidence interval
              color = "black",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) + 
  theme_light() + 
  ggpubr::stat_regline_equation(label.x = 1200, label.y = 90) + # Need to adjust trendline so that graph isn't wonky
  labs(x = "Weight (g)", y = "Hind foot length (mm)") 
```

**Figure 5:** Relationship between weight (g) and hind foot length (mm) for juvenile snowshoe hares. Linear model summary: $\beta$~1~ = `r round(hares_lm_tidy$estimate[2], 2)` g mm^-1^, p < 0.001, R^2^ = `r round(hares_lm_glance$r.squared,2)`, Pearson's *r* = `r round(hares_cor_tidy$estimate,2)`). Data: Kielland et al. (2017).

## Summary

Findings revealed through exploratory data analysis:

- Number of juvenile snowshoe hares trapped was greatest during the first two years of the study (1999 and 2000). Trap counts fluctuated throughout the remaining years, with higher counts recorded roughly every three years. 
- Initial data visualization revealed that male juvenile hare weights were higher on average than female weights. The weights of juvenile hares caught at the Bonanza Black Spruce site  were higher on average than individuals trapped at Bonanza Mature and Bonanza Riparian. 
- The difference in means of male and female juvenile hare weights is significant, although the effect size is small. 
- There is a moderate positive linear correlation between juvenile hare weight (g) and hind foot length (mm), although outliers present suggest a slight concern in the validity of this model. 

Potential next steps:

- Explore correlation between vegetation density of sites and hare weight. Why does the Bonanza Black Spruce site have higher average juvenile snowshoe hare weights? Perhaps this site offers more protection from predators than others. 
- Explore potential models that might fit the weight versus foot length data more accurately. 
- Explore relationship between juvenile hare weight and ear length.
- Explore data for adult snowshoe hares. 

## Citations 

[Feierabend, D. and Kielland, K. 2017. Movements, activity patterns, and habitat use of snowshoe hares (*Lepus americanus*) in interior Alaska. Journal of Mammology vol 95. ]( https://academic.oup.com/jmammal/article/95/3/525/877149)

[Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. ](https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14) 