---
title: "Mono Lake vizualization"
description: |
  exploring Mono Lake water levels over time
author:
  - name: Grace Kumaishi
date: 12-04-2020
output:
  distill::distill_article:
    self_contained: false
    code_folding: Code
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(plotly)
library(ggplot2)
```

```{r}
# Read in the data

mono_lake <- read.csv(here("data", "mono_lake_annual_levels.csv"), skip = 5) %>% 
  clean_names() %>% 
  rename(lake_level_ft = lake_level_feet_above_sea_level)
```

### Assignment: create a beautiful visualization: 
```{r, fig.align = "center", fig.width = 9}
# Create a beautiful data visualization of lake levels 

mono_lake_levels <- #store ggplot to make plot interactive

ggplot(mono_lake) +
  geom_line(aes(x = year, 
                y = lake_level_ft), 
            color = "skyblue4", 
            size = .5) +
  geom_ribbon(aes(x = year, 
                  y = lake_level_ft, 
                  xmin = 1850, 
                  xmax = 2017, 
                  ymin = 6340, 
                  ymax = lake_level_ft), 
              fill = "skyblue2", 
              alpha = .5) +
  labs(y = "Lake level (ft)",
       x = "Year") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = c(1850, 1875, 1900, 1925, 1950, 1975, 2000)) +
  geom_segment(aes(x = 1850,
                   xend = 2017,
                   y = 6377,
                   yend = 6377), 
             color = "darkorange", 
             size = .3) +
  geom_segment(aes(x = 1850,
                   xend = 2017,
                   y = 6360,
                   yend = 6360), 
             color = "darkorange2", 
             size = .3) +
  geom_segment(aes(x = 1850,
                   xend = 2017,
                   y = 6350,
                   yend = 6350), 
             color = "darkorange4", 
             size = .3) +
  annotate("text", 
           x = 1878, 
           y = 6379, 
           label = "Land bridge emerges (6,377 ft)", 
           color = "black", 
           size = 3, 
           fontface = "bold") +
  annotate("text", 
           x = 1881, 
           y = 6362, 
           label = "Salinity reaches 120 g/L (6,360 ft)", 
           color = "black", 
           size = 3, 
           fontface = "bold") +
  annotate("text", 
           x = 1881, 
           y = 6352, 
           label = "Salinity reaches 150 g/L (6,350 ft)", 
           color = "black", 
           size = 3, 
           fontface = "bold") +
  annotate("point", 
           x = c(1941, 1978, 1983, 1994),
           y = c(6417, 6375, 6378.6, 6374.6),
           alpha = 0.8,
           color = "gray25") +
  annotate("segment",
           x = c(1941, 1978, 1983, 1994),
           xend = c(1941, 1978, 1983, 1994),
           y = c(6425, 6383, 6388.6, 6384.6),
           yend = c(6417, 6375, 6378.6, 6374.6),
           color = "gray25",
           size = .2) +
  annotate("text",
           x = 1941,
           y = 6427,
           label = "1941",
           color = "black",
           size = 3) +
  annotate("text",
           x = 1978,
           y = 6385,
           label = "1978",
           color = "black",
           size = 3) +
  annotate("text",
           x = 1983,
           y = 6390.6,
           label = "1983",
           color = "black",
           size = 3) +
  annotate("text",
           x = 1994,
           y = 6386.6,
           label = "1994",
           color = "black",
           size = 3) 

ggplotly(mono_lake_levels) #This will make the graph interactive, works in knitted doc!! 
```

**Figure 1:** Annual lake level measurements (feet above sea level) between 1850 and 2017 collected from the [Mono Basin Clearinghouse](https://www.monobasinresearch.org/). Lake levels that pose significance to nesting bird populations are noted with horizontal lines. At 6,377 ft above sea level, a land bridge emerges that allows predators access to crucial California Gull nesting habitat. Salinity reaches 120 g/L at 6,360 ft above sea level and 150 g/L at 6,350 feet above sea level, resulting in decreased food availability for Eared Grebes and California Gulls. Notable years are annotated on the graph (1941: LADWP began diverting water from Mono Lake's tributary streams; 1978: Mono Lake Committee formed; 1983: California Supreme Court decision; 1994: State Water Board decision). Further information about these dates and more can be found on the Mono Lake [website](https://www.monolake.org/learn/aboutmonolake/savingmonolake/#water-diversions-almost-destroyed-mono-lake).  Data: National Research Council (1987).

### Citation:
[National Research Council. 1987. The Mono Basin Ecosystem: Effects of Changing Lake Level. Washington, DC: The National Academies Press. https://doi.org/10.17226/1007.](https://www.nap.edu/catalog/1007/the-mono-basin-ecosystem-effects-of-changing-lake-level)
